---
editor_options: 
  markdown: 
    wrap: 72
---

|                                        |
|:--------------------------------------:|
| R code for GV900 Homework Assignment 2 |
|          University of Essex           |

### Task 1- Load 6 r packages

```{r warning=FALSE}
library(ggplot2)
library(gmodels)
library(Hmisc)
library(stargazer)
library(effects)
library(gridExtra)
```

### Task 2- Load Titanic csv file and store as object called td.

```{r}
td.fp <- file.path("data","titanic.csv")
td. <- read.csv(td.fp)
```

### Task 3- How many passengers were on board?

```{r}
nrow(td.) #Answer: 1309
```

### Task 4 - Create a frequency table for the survived variable

```{r}
table(td.$survived)
```

### Task 5- Calculate the survival percentage

1309 passengers

500 survived

500/1309\*100 = 38.1971%

### Task 6- Create a frequency table for class tickets

```{r}
table(td.$pclass)
```

### Task 7- Identifying Variables

**What is the dependent variable of the investigation?**

*Answer: Passenger Survival*

**What is the independent variable of the investigation?**

*Answer*: *Socio-Economic class of passengers*

### Task 8- Create a cross-tabulation table for 'pclass' and 'survived' variables

```{r}
CrossTable(td.$ survived, td.$ pclass, prop.r = FALSE,
prop.t = FALSE,
prop.chisq = FALSE, chisq = TRUE)
```

### Task 9 - Interpret the cross-tabulation table

a)  **What is the survival rate among 1st class passengers**

*Answer a): 61.9%*

b)  **What is the survival rate among 2nd class passengers**

    *Answer b): 43.0%*

c)  **What is the survival rate among 3rd class passengers**

    *Answer c): 25.5%*

### Task 10- Relationship with socio-economic class and survival

According to the statistics, there is a negative relationship between
the passengers socio-economic class, and the rate of survival. It
appears that the richer the passengers were, the higher likelihood that
they would survive the sinking, as we predicted in our hypothesis.

### Task 11- Summarizing the results

*Answer 1): j (0.05)*

*Answer 2): n (reject)*

*Answer 3): g (95%)*

*Answer 4): p (do not find)*

### Task 12- Explore the relationship between female passengers and fare

```{r}
td.Fare <- td.[!is.na(td.$fare),]
```

```{r message=FALSE}
g <- ggplot(td.Fare, aes(x = fare)) + geom_histogram()
g <- g + theme(axis.text.x = element_text(size = 14))
g <- g + xlab("Price of Ticket") + ylab("Gender of Passengers")
g <- g + stat_bin(bins=50)
g <- g + facet_grid(~ female)
g
```

```{r}
by(data = td.Fare $ fare,  
   INDICES = td.Fare $ female,   
   FUN = summary)             
```

```{r}
t.test( fare ~ female,      
        data = td.Fare,   
        var.equal = TRUE) 
```

### Task 13- Interpret the results of the bivariate test

Looking at the graph, it appears that females pay more for their ticket
then men do, because the vast majority of the male tickets were priced
under at £100 or under, whereas there is still a significant amount of
females who paid over that amount.

This appears true of we look at the mean price comparison. The mean
price for females was £46.20 whereas the mean price for men was £26.15.

Overall, looking at the graphs and statistics, it is evident that female
passengers had more expensive tickets compared to men

### Task14- Explore the relationship between age and fare graphically

```{r}
# Remove rows with NULL age
td.age <- td.[!is.na(td.$age),]
```

```{r warning = FALSE}
g <- ggplot(td.age, aes( x=age, y=fare) ) + geom_point()
g <- g + ylab("Price of Ticket") + xlab("Age of Passengers")
g
```

### Task 15- Perform an appropriate bivariate statistical test to explore the relationship between age and fare

```{r}
td.ag.fr <- na.omit(td.[,c(4,11)])

lab.uni.mat <- as.matrix(td.ag.fr[ c("fare","age")])
rcorr(lab.uni.mat, type = "pearson")
        
```

### Task 16- Interpret the results of the bivariate test

Looking at the graph, there doesn't appear to any correlation between
age and fare.

This appears to be true in the statistics, as the p value = 0.

This means we need to reject the null hypothesis To conclude, there
isn't a significant relationship between age and fare

### Task 17- Regress fare on age, and produce a regression table using the stargazer function.

```{r}
fit.1 <- lm(fare ~ age, data = td.)

stargazer(fit.1, type = "text")
```

### Task 18- Create a plot that illustrates the estimated effect of age on fare based on the model you estimated above

```{r}
effect("age", mod = fit.1)
eff.age <- effect("age", mod = fit.1)
plot(eff.age)

plot(eff.age, 
     main = "Affect of Age on Ticket Prices",
     xlab = "Age of Passengers",
     ylab = "Ticket Fares")
```

### Task 19- Looking at the graph and the regression table, is there a positive correlation between age and fare

Looking at the graphical results, it is evident that there is a positive
correlation between age and fare prices. Statistically, we can see that
this is true because the R2 number is relatively high at 0.032. To
conclude, we can see that on average, the older the passenger is, the
higher the likelihood that the passengers pays a higher fare.

### Task 20- Graphically explore the relationship between age and fare, holding constant the female variable

```{r}
td.ag.ge.fr <- na.omit(td.[,c(4,7,11)])

g <- ggplot(td.ag.ge.fr, aes( x=age, y=fare) ) + geom_point()
g <- g + ylab("Price of Ticket") + xlab("Age of Passengers")
g <- g + facet_grid(~ female)
g
```

### Task 21

### Using stargazer, regress fare on age and female

```{r}
# New Model fare ~ female
fit.2 <- lm(fare ~ age + female, data = td.)

# Summarise model
summary(fit.2)

```

Compare the model from Task 17 (fit.1) with the model from this Task
(fit.2)

```{r}
 suppressWarnings( # Using supressWarnings here to remove length NULL message
                  stargazer(fit.1,fit.2, type = "text") 
                  )

```

### Task 22

### Which one of the two regression models performs better?

*Answer 1): D (Adjusted R2)*

*Answer 2): F (Smaller)*

*Answer 3): H (First)*

*Answer 4): I (Second*)

### Task 23 - Estimate the effect of female on fare based on the second regression model you estimated

```{r}
eff.fr.ge <- effect(term = "female", 
                      mod = fit.2)
(plot.fr.ge <- plot(eff.fr.ge,  main = "Gender Effect Plot", ylim = c(0, 110)))

```

### Task 24 - Create a plot that estimates age on fare for male and female passengers separately based on the second regression model

#### Effects and Plot solution (holding Female variable constant)

Convert female to binary variable. Run linear regression using new binary variable. Generate effect objects for
Male and Female holding on female as constant.
```{r}
female.1 <- ifelse(td. $ female == "Female", 1, 0)
fit.3 <- lm(fare ~ age + female.1, data = td.)
eff.fr.ag.ge.1 <- effect(term = "age", 
                      mod = fit.3, 
                      given.values = c("female.1" = 1))
eff.fr.ag.ge.0 <- effect(term = "age", 
                      mod = fit.3, 
                      given.values = c("female.1" = 0))
```

Generate plot objects.
```{r}
plot.fr.ag.ge.1 <- plot(eff.fr.ag.ge.0,  main = "Female Effect Plot"  , ylim = c(0, 110))
plot.fr.ag.ge.0 <- plot(eff.fr.ag.ge.0,  main = "Male Effect Plot"  , ylim = c(0, 110))

```

Output plots in 2 column, side-by-side, for female and male.
```{r}
grid.arrange(plot.fr.ag.ge.1, plot.fr.ag.ge.0, ncol = 2)
```

**End of File**
